---
title: 'Tidy Tuesday #1'
author: "Erin Franke"
date: "5/24/2021"
output:
  html_document:
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

### Mario Kart 64 Tidy Tuesday Plot

Load packages and the Mario Kart data from Github. 
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidytuesdayR)
library(tidyverse)
library(ggthemes)
records <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')
```

**Question: "How much faster is the single lap record (multiplied 3x) than the 3 lap course record for each of the 16 Mario Kart 64 courses? Specifically, is this difference greater for longer courses than shorter courses?"** 

Create table of three lap records, filtering out any records made using a shortcut. 
```{r}
(threelaprecords <- records %>%
  filter(type == "Three Lap", shortcut == "No") %>%
  group_by(track) %>%
  summarize(minTime = min(time, na.rm = TRUE)) %>%
  arrange(desc(minTime)))
```

Create table of one lap records, and multiply by three. Again filter out any records made with a shortcut. 
```{r}
onelaprecords <- records %>%
  filter(type == "Single Lap", shortcut == "No") %>%
  group_by(track) %>%
  summarize(singlelap = min(time, na.rm = TRUE)) %>%
  arrange(desc(singlelap))

(records1lap3x <- onelaprecords %>%
  mutate(singlelap3x = singlelap * 3) %>%
  select(track, singlelap3x))
```

Join the tables and create a column called "difference" which is how much faster the single lap record (multiplied 3x) is than the three lap course record. Additionally, I created a column called "Course Length", which denotes a course as short, medium, or long based on the three lap record. \
Long courses: Three Lap Record greater than 170 seconds \
Medium courses: Three Lap Record between 115-135 seconds \
Short courses: Three Lap Record less than 105 seconds\
```{r}
differences <- threelaprecords %>%
  full_join(records1lap3x, by = "track")%>%
  rename(`Three Lap Record` = `minTime`) %>%
  rename(`One Lap Record (3x)` = `singlelap3x`) %>%
  mutate(difference = `Three Lap Record` - `One Lap Record (3x)`) %>%
  mutate(`Course Length` = `Three Lap Record`)

differences$`Course Length` <- cut(differences$`Course Length`, 
                   breaks=c(-Inf, 115, 133, Inf), 
                   labels=c("short (<105 seconds)","medium (115-135 seconds)","long (>170 seconds)"))
differences
```

Create the bar plot, coloring each bar by course length. 
```{r}
differences %>%
  ggplot(aes(x=difference, y=fct_reorder(track, difference), fill = `Course Length`))+
  geom_col(color = "white") + 
  theme_minimal() + 
  labs(title = "Comparing course records for Mario Kart 64", subtitle = "How much faster is the 1 lap record (multplied 3x) than the 3 lap course record?", y = "", caption = "Plot created by Erin Franke, data from Mario Kart World Records") + 
  scale_x_continuous(expand = c(0,0))+
  scale_fill_wsj()
```

### Plot made with group in class: 

**Research question: Which track is the fastest during a three lap, no shortcuts race?**

Filter the data
```{r}
records_filtered <- records %>%
  filter(type == "Three Lap", shortcut == "No") %>%
  group_by(track, system_played) %>%
  summarize(recordtime = min(time, na.rm=TRUE))
```

Create plot
```{r}
records_filtered %>%
  ggplot(mapping = aes(x=recordtime, y=fct_reorder(track,recordtime), fill=system_played)) +
  geom_col(position = "dodge") + 
  labs(x= "Record Time", fill = "System Played", y = "", title = "Record Time by Track for Mario Kart 64", subtitle = "For three lap and no shortcut races", caption = "Plot created by Erin Franke, data from Mario Kart World Records")+
  scale_x_continuous(expand = c(0,0))+
  theme_minimal()
```

