---
title: "Tidy Tuesday Week 3 - Chocolate Bars"
author: "Erin Franke"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(tidyverse)
library(tidyr)
library(ggtext)
library(showtext)
font_add_google("Gochi Hand", "gochi")
font_add_google("Schoolbell", "bell")
font_add_google("Kanit", "kanit", bold.wt = 600)
showtext_auto()
chocolate <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')
```

Get a sense of what types of categories we have:
```{r}
# 580 manufacturers
chocolate %>%
  group_by(company_manufacturer) %>%
  count()
# 67 locations
chocolate %>%
  group_by(company_location) %>%
  count() %>%
  arrange(desc(n))
# 62 countries of origin
chocolate %>%
  group_by(country_of_bean_origin) %>%
  count() %>%
  arrange(desc(n))
# ingredients 
chocolate %>%
  group_by(ingredients) %>%
  count()
# ref
chocolate %>%
  group_by(ref) %>%
  count()
```

Try to understand the `most_memorable_characteristics` column! 
```{r}
chocolate %>%
  mutate(id = row_number(), 
    most_memorable_characteristics = strsplit(as.character(most_memorable_characteristics), ",")) %>%
  unnest(most_memorable_characteristics) %>%
  pivot_wider(id_cols = c(id, rating, ingredients, cocoa_percent, specific_bean_origin_or_bar_name, country_of_bean_origin, review_date, company_location, company_manufacturer, ref), names_from = most_memorable_characteristics, values_from = most_memorable_characteristics) %>%
  summarise_all(funs(sum(is.na(.)))) %>%
  pivot_longer(cols = -c(id:ref), names_to = "characteristic", values_to = "num_na") %>%
  mutate(times = 2530 - num_na) %>%
  arrange(desc(times)) %>%
  select(characteristic, times)
```

Create additional columns for top 8 most used characteristics and each ingredient.
```{r}
clean_choc <- chocolate %>%
  mutate(cocoa_percent = as.numeric(substr(cocoa_percent, 1, nchar(cocoa_percent) -1)), 
         num_ingredients  = as.numeric(substr(ingredients, 1, 1)), 
         beans = case_when(as.integer(str_detect(ingredients,"B")) == 1 ~ 1, 
                           TRUE ~ 0), 
         sugar = case_when(as.integer(str_detect(ingredients,"S")) == 1 ~ 1, 
                           TRUE ~ 0), 
         cocoa_butter = case_when(as.integer(str_detect(ingredients,"C")) == 1 ~ 1,
                                  TRUE ~ 0), 
         vanilla = case_when(as.integer(str_detect(ingredients,"V")) == 1 ~ 1,
                                  TRUE ~ 0), 
         letchin = case_when(as.integer(str_detect(ingredients,"L")) == 1 ~ 1,
                                  TRUE ~ 0), 
         salt = case_when(as.integer(str_detect(ingredients,"Sa")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_cocoa = case_when(as.integer(str_detect(most_memorable_characteristics, "cocoa")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_sweet = case_when(as.integer(str_detect(most_memorable_characteristics, "sweet")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_nutty = case_when(as.integer(str_detect(most_memorable_characteristics, "nutty")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_creamy = case_when(as.integer(str_detect(most_memorable_characteristics, "creamy")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_roasty = case_when(as.integer(str_detect(most_memorable_characteristics, "roasty")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_earthy = case_when(as.integer(str_detect(most_memorable_characteristics, "earthy")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_sandy = case_when(as.integer(str_detect(most_memorable_characteristics, "sandy")) == 1 ~ 1,
                                  TRUE ~ 0), 
         char_sour = case_when(as.integer(str_detect(most_memorable_characteristics, "sour")) == 1 ~ 1,
                                  TRUE ~ 0))
```

Cluster the chocolates using several variables. On the final plot, plot the drinks by `cocoa_percent` and `rating`. 
```{r}
choc_sub <- clean_choc %>%
  filter(num_ingredients >=1) %>%
  select(num_ingredients, vanilla, char_sandy, char_creamy, letchin, salt, char_sweet, char_nutty, char_roasty, char_sour)
```

```{r}
# Run k-means for k = centers = 3
set.seed(253)
kclust_k2 <- kmeans(choc_sub, centers = 2)
kclust_k3 <- kmeans(choc_sub, centers = 3)
kclust_k4 <- kmeans(choc_sub, centers = 4)
```

```{r}
# Run k-means on the *scaled* data (all variables have SD = 1)
set.seed(253)
kclust_k2_scale <- kmeans(scale(choc_sub), centers = 2)
kclust_k3_scale <- kmeans(scale(choc_sub), centers = 3)
kclust_k4_scale <- kmeans(scale(choc_sub), centers = 4)
clean_choc <- clean_choc %>%
  filter(num_ingredients >=1) %>%
    mutate(kclust_3_scale = factor(kclust_k3_scale$cluster), 
           kclust_4_scale = factor(kclust_k4_scale$cluster), 
           kclust_2_scale = factor(kclust_k2_scale$cluster))
```

Check out the within-cluster sum of squares.
```{r}
choc_cluster_ss <- function(k){
    # Perform clustering
    kclust <- kmeans(scale(choc_sub), centers = k)

    # Return the total within-cluster sum of squares
    return(kclust$tot.withinss)
}

tibble(
    k = 1:15,
    tot_wc_ss = purrr::map_dbl(1:15, choc_cluster_ss)
) %>% 
    ggplot(aes(x = k, y = tot_wc_ss)) +
    geom_point() + 
    labs(x = "Number of clusters",y = 'Total within-cluster sum of squares') + 
    theme_classic()
```

Create final graph. 
```{r}
clean_choc %>%
    group_by(kclust_3_scale) %>%
    summarize(across(c(rating, num_ingredients, char_earthy, char_nutty, char_roasty, char_sour, char_sandy, salt, letchin, cocoa_butter, vanilla, char_creamy), mean))

ggplot(clean_choc, aes(x = cocoa_percent, y = rating, color = kclust_3_scale)) +
  geom_jitter()+
  labs(title = "We can't learn too much about chocolate bars when clustering on their ingredients and memorable characteristics", y = "rating 1-5", x = "cocoa percent", caption = "Data courtesy of Flavors of Cacao")+
  theme_minimal()+
  theme(plot.title = element_markdown(family = "bell", face = "bold", color = "white", size = 9),
        plot.title.position = "plot",
        axis.text = element_markdown(family = "bell", color = "white"), 
        axis.title =  element_markdown(family = "bell", color = "white"), 
        axis.ticks = element_blank(), 
        axis.line = element_line(color = "white"),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#5F3E3D"), 
        legend.position = "none", 
        plot.caption = element_markdown(family = "bell", color = "white"))+
  scale_color_manual(values = c("#F24A07", "#FBD900", "black"))+
  annotate(geom="text", y=1.2, x=90, label = "avg 3.08 ingredients & a 3.23 \navg rating", family = "gochi", cex=3, color = "black")+
  annotate(geom="text", y=1.5, x=53, label = "all described as sour but maintain an \navg rating of 3.03 with 2.76 ingredients \non average", family = "gochi", cex=3, color = "#FBD900")+
  annotate(geom="text", y=3.2, x=50, label = "3.22 avg rating with 2.89 \ningredients on avg, \nall described as roasty", family = "gochi", cex=3, color = "#F24A07")
```

