---
title: "Try to analyze Wordle Data?"
author: "Erin Franke"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(tidyverse)
library(janitor)
library(ggtext)
library(showtext)
font_add_google("Gochi Hand", "gochi")
font_add_google("Schoolbell", "bell")
font_add_google("Kanit", "kanit", bold.wt = 600)
showtext_auto()

twirdle <- function(tweets) {
  
  g   <- "\U1F7E9"
  o   <- "\U1F7E7"
  y   <- "\U1F7E8"
  blu <- "\U1F7E6"
  bla <- "\U2B1B"
  w   <- "\U2B1C"
  
  rx_all   <- paste(g, o, y, blu, bla, w, sep = "|")
  rx_right <- paste(g, o, sep = "|")
  rx_place <- paste(y, blu, sep = "|")
  rx_wrong <- paste(bla, w, sep = "|")
  rx_color <- paste(o, blu, sep = "|")

  tweets$meta <- regexpr(
    "Wordle \\d{1,} [\\d{1}|X]/\\d{1}",
    tweets$text,
    perl = TRUE
  )
  
  tweets$meta <- setNames(
    tweets$meta, 
    ifelse(tweets$meta < 0, FALSE, TRUE)
  )
  
  tweets$meta <- ifelse(
    names(tweets$meta),
    regmatches(tweets$text, tweets$meta),
    NA_character_
  )
  
  tweets <- tweets[!is.na(tweets$meta), ]
  
  tweets$edition <- as.numeric(
    regmatches(
      tweets$meta, 
      regexpr("\\d{1,}", tweets$meta)
    )
  )
  
  tweets$attempts <- regmatches(
    tweets$meta,
    regexpr("[\\d{1}|X](?=/)", tweets$meta, perl = TRUE)
  )
  
  tweets$attempts <- ifelse(
    tweets$attempts == "X",
    NA_character_,
    tweets$attempts
  )
  
  tweets$attempts <- as.numeric(tweets$attempts)
  
  tweets$allowed <- as.numeric(
    regmatches(
      tweets$meta, 
      regexpr("(?<=/)\\d{1}", tweets$meta, perl = TRUE)
    )
  )
  
  tweets$grid <- regmatches(
    tweets$text, 
    gregexpr(rx_all, tweets$text) 
  )
  
  tweets$grid <- lapply(
    tweets$grid, 
    function(x) paste(x, collapse = "")
  )
  
  tweets$colorblind <- ifelse(
    grepl(rx_color, tweets$grid), TRUE, FALSE
  )
  
  tweets$mode <- ifelse(
    grepl(bla, tweets$grid), "dark",
    ifelse(grepl(w, tweets$grid), "light", "unknown")
  )
  
  tweets$grid <- gsub(rx_right, "G", tweets$grid)
  tweets$grid <- gsub(rx_place, "Y", tweets$grid)
  tweets$grid <- gsub(rx_wrong, "-", tweets$grid)
  
  tweets[, c("edition", "attempts", "allowed", "mode",
             "colorblind", "grid", "status_id")]
  
}
```

Load in tweets
```{r, eval=FALSE}
tweets <- rtweet::search_tweets(
  q = "Wordle \U2B1B OR \U2B1C",
  n = 10000,  # return 10 tweets
  include_rts = FALSE  # no retweets
)
```

Filter for twirdle edition 213 and remove any attempts that were 0. Create a first guess and last guess column, and whether the final guess was correct. 
```{r}
twirdle213 <- twirdle(tweets) %>%
  filter(edition == 213, attempts >= 1) %>%
  mutate(first_guess = substr(grid, 1, 5), 
         last_guess = substr(grid, nchar(grid)-4, nchar(grid)), 
         correct = case_when(last_guess == "GGGGG" ~ "yes",
                             last_guess == "" ~ "yes",
                             TRUE ~ "no"), 
         attempts2 = as.integer(nchar(grid)/5))
```

Learn some stats about wordle 213:
```{r}
#learn some stats about wordle 213 - 95.7% of people were about to complete the wordle
twirdle213 %>%
  group_by(correct) %>%
  count()
5649/(5649+253)

# for those that completed the wordle, the average number of attempts was 4.308
twirdle213 %>%
  filter(correct == "yes") %>%
  summarize(avg = mean(attempts2))

# find how many people did not get any letters correct the first word: 22.04% of people got no letters correct in their first word
twirdle213 %>%
  group_by(first_guess) %>%
  count()
1301/5902

# find how many people got at least one green on the first try - 40.36 %
twirdle213 %>%
  mutate(green = case_when(as.integer(str_detect(first_guess,"G")) == 1 ~ 1, 
                           TRUE ~ 0)) %>%
  group_by(green) %>%
  count()
2382/(2382+3520)

# for those that it took at least 4 attempts, the average number of Green letters in word 3 was 2.025
twirdle213 %>%
  filter(attempts2 >=4) %>%
  mutate(guess3 = substr(grid, 11, 15), 
         count3G = str_count(guess3, "G")) %>%
  summarize(mean(count3G))

# no one got it in 1 attempt
twirdle213 %>%
  filter(attempts2 == 1)
```


Looking at Wordle 213: Remove people whose last guess was incorrect and display the number of tries it took for people to get the wordle who did not give up and got it in 6 or fewer tries. 
```{r}
twirdle213 %>%
  filter(correct == "yes", 
         attempts2 >0 & attempts2 <=6) %>%
  tabyl(attempts2) %>%
  ggplot(aes(x=attempts2, y=percent))+
  geom_col(fill = "#6DA962")+
  theme_minimal()+
  labs(title = "A quick analysis of Wordle 213 from 5,902 tweets - no one guessed it on attempt 1!", y = "proportion of users", x = "attempts", caption = "Courtesy to Matt Dray for twirdle function!")+
  geom_vline(xintercept = 4.308, color = "white", linetype = "dashed")+
  ylim(0, 0.3)+
  xlim(1, 6)+
  theme_minimal()+
  theme(plot.title = element_markdown(family = "bell", face = "bold", color = "white", size = 11),
        plot.title.position = "plot",
        axis.text = element_markdown(family = "bell", color = "white"), 
        axis.title =  element_markdown(family = "bell", color = "white"), 
        axis.ticks = element_blank(), 
        axis.line = element_line(color = "white"),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#7B7B7B"), 
        legend.position = "none", 
        plot.caption = element_markdown(family = "bell", color = "white"))+
  annotate(geom="text", y=0.25, x=1.5, label = "Fact 1: 95.7% of tweeters \ncompleted this wordle", family = "bell", cex=3.2, color = "#CCB35B")+
  annotate(geom="text", y=0.2, x=4.31, label = "Avg number of \nattempts was 4.308", family = "bell", cex=3.2, color = "white")+
  annotate(geom="text", y=0.17, x=1.5, label = "Fact 2: 22.04% of tweeters \n got no letters correct (all \ngray) on their 1st attempt", family = "bell", cex=3.2, color = "#CCB35B")+
  annotate(geom="text", y=0.1, x=1.6, label = "Fact 3: 40.36% of tweeters got \n1+ green letters on attempt 1", family = "bell", cex=3.2, color = "#CCB35B")+
  annotate(geom="text", y=0.265, x=3, label = "Fact 4: For those \nit took 4+ attempts, \nthe average number \nof green letters in \nword 3 was 2.025", family = "bell", cex=3.2, color = "#CCB35B")
```


